# Environmental Justice Policy 2010 - 2016

The de facto Massachusetts environmental justice policy for 2010-2016 defined environmental justice communities as neighborhoods (U.S. Census Bureau census block groups) that meet *one or more* of the following criteria: 

* The median annual household income is at or below 65 percent of the statewide median income for Massachusetts; *or*
* 25 percent of the residents are minority; *or*
* 25 percent of the residents are lacking English language proficiency.

Using this definition and the American Community Survey 5-year Estimates for 2013-2017 for block groups in Massachusetts, we find that `r {round(sum(blkgrp_shp_WGS84$CRITCNT2010 > 0) / nrow(blkgrp_shp_WGS84) * 100,1)}`% of Massachusetts block groups are classified as environmental justice communities. This classification can be visualized in Figure \@ref(fig:treemap2010) below, which shows the percentage of block groups that fall into each category or combination of environmental justice criteria.

```{r treemap2010, echo=FALSE, message=FALSE, warning=FALSE, fig.align = "center", fig.cap="Tree map of block groups classified as environmental justice by 2010-2016 policy."}
# Create a tree map with treemapify
# library(treemapify)
# First, you need to generate a summary df with one value per observation or EJ criterion
blkgrp_shp_WGS84 %>% 
  mutate(EJCRIT2010 = case_when(EJCRIT2010 == "E" ~ "English Isolation",
                                    EJCRIT2010 == "E  M" ~ "Minority and English Isolation", 
                                    EJCRIT2010 == "E I" ~ "Income and English Isolation",
                                    EJCRIT2010 == "E I M" ~ "Minority, Income, and English Isolation",
                                    EJCRIT2010 == "I" ~ "Income",
                                    EJCRIT2010 == "I M" ~ "Minority and Income",
                                    EJCRIT2010 == "M" ~ "Minority",
                                    EJCRIT2010 == "" ~ "Not EJ")) %>% 
  group_by(EJCRIT2010) %>% 
  summarize(blkgrps = n()) %>% 
  mutate(percent = blkgrps/sum(blkgrps)*100) %>% 
  ggplot(aes(area = percent, fill=EJCRIT2010, label = paste(EJCRIT2010,paste(round(percent,1),"%"),sep="\n"))) + 
  scale_fill_manual(name = "Environmental Justice Criteria", 
                    values = c("lightskyblue","lawngreen","cyan","gold","deeppink","darkgreen","navyblue","gray88")) +
  labs(title = "Massachusetts Census Blockgroups Meeting 2010 Environmental Justice Criteria",
       caption = "Based on 2010 MA EJ Policy and ACS 5yr 2017 Blockgroup Estimates") +
  geom_treemap() + 
  geom_treemap_text(color = "white", place = "centre", grow = TRUE)
```

These classifications can be visualized and explored geographically in the interactive map below. Click on the search tool (white box in upper left of map) and type in a town or city name to zoom to that municipality. Click on individual block groups to see more detailed demographics. 

```{r map2010, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Map of block groups classified as environmental justice by 2010-2016 policy."}
# Create map of 2010 EJ policy polygons across state
# create df without "" in EJCRIT2010 for setting up color palette
blkgrp_shp_WGS84_naEJCRIT2010 <- blkgrp_shp_WGS84 %>%
  select(NAME,TOWN,EJCRIT2010,pctminority,medhhincE,pctlimitenghh) %>% 
  filter(EJCRIT2010 !="")

# How many EJ blockgroups within top quintile communities?
inside2010 <- blkgrp_shp_WGS84_naEJCRIT2010 %>% 
  as.data.frame() %>% 
  group_by(TOWN) %>% 
  summarize(inside = n()) %>% 
  left_join(macosub_shp_WGS84,.,by=c("TOWN" = "TOWN")) %>% 
  na.omit()

# set up color palette with colors in order of criteria, which are in alphabetical order by default
# sort(unique(blkgrp_shp_WGS84_naEJCRIT2010EJCRIT2010$EJCRIT2010))# confirm order of criteria
EJ2010pal <- colorFactor(palette = c("lightskyblue","deeppink","cyan","navyblue","lawngreen","darkgreen","gold"), 
                         domain = blkgrp_shp_WGS84_naEJCRIT2010$EJCRIT2010)

# create vector colors in order to be used in legend
EJcols <- c("gold","lawngreen","lightskyblue","darkgreen","deeppink","cyan","navyblue")
# convert named colors to hexadecimal
EJcolsHex <- col2hex(EJcols)

# Create popup. NOTE THAT POPUP AND PALETTE MUST BE CREATED FROM EXACT SAME DF AS USED IN leaflet() MAP. OTHERWISE, POPUPS WILL SHOW UP IN THE WRONG LOCATION!
EJ2010PopUp <- paste0(blkgrp_shp_WGS84_naEJCRIT2010$NAME, "<br/>",
                      "<b>Town:</b> ", blkgrp_shp_WGS84_naEJCRIT2010$TOWN, "<br/>",
                      "<b>2010 EJ Criteria Met:</b> ", blkgrp_shp_WGS84_naEJCRIT2010$EJCRIT2010, "<br/>",
                      "<b>ACS 5yr 2017</b> ", "<br/>",
                      "<b>% Minority: </b>", round(blkgrp_shp_WGS84_naEJCRIT2010$pctminority,1),"<br/>",
                      "<b>Median Household Income: </b>", paste0("$",round(blkgrp_shp_WGS84_naEJCRIT2010$medhhincE)),"<br/>",
                      "<b>% Limited English Households: </b>", round(blkgrp_shp_WGS84_naEJCRIT2010$pctlimitenghh,1))

# draw the map of EJ Polygons according to 2010 policy
leaflet() %>% 
  addProviderTiles(providers$Stamen.TonerLite) %>% 
  addPolygons(data = town_shp_WGS84,
              weight = 0.7,
              opacity = 1,
              color = "gray",
              fillOpacity = 0,
              label=~TOWN, popup=~TOWN, group='muni') %>% 
  addPolygons(data = blkgrp_shp_WGS84_naEJCRIT2010,
              fillColor = ~EJ2010pal(EJCRIT2010), 
              weight = 0.5,
              opacity = 0.7,
              color = "white",
              dashArray = 3,
              fillOpacity = 0.7,
              label=~TOWN, 
              popup = EJ2010PopUp,
              highlight = highlightOptions(
                weight = 5,
                color = "#666",
                dashArray = "",
                fillOpacity = 0.7,
                bringToFront = TRUE)) %>% 
  addLegend(colors = EJcolsHex, 
            labels = c("Minority","Income","English Isolation","Minority and Income","Minority and English Isolation","Income and English Isolation","Minority, Income, and English Isolation"), title = "Environmental Justice Populations", position = "bottomleft") %>% 
  setView(lng = -72, 42.1, zoom = 8) %>% 
  # addMiniMap() %>% 
  addScaleBar(position = "bottomright") %>% 
  addSearchFeatures(targetGroups  = 'muni', 
                    options = searchFeaturesOptions(zoom=14, openPopup=TRUE, hideMarkerOnCollapse=T))
```


