# Environmental Justice Alternative Policy D - Modify Minority and English Isolation Criteria by Income AND Education

This option adds two modifiers to the criteria for minority and English isolation. Specifially, block groups qualify for EJ status by minority or English isolation only if the median income of the block group does not exceed 125% of the statewide median AND no more than half of the population 25 years and over has a college degree. This means that, regardless of criteria, the median income cannot exceed \$92,708.75 and the percentage of the population 25 years and older with a 4-year college degree or higher cannot exceed 50%. 

Under Alternative Environmental Justice Policy D option, environmental justice communities are defined as neighborhoods (U.S. Census Bureau census block groups) that meet *one or more* of the following criteria: 

* The median annual household income is at or below 65 percent of the statewide median income for Massachusetts; *or*
* 25 percent or more of the residents are minority *AND the median annual household income does not exceed 125 percent of the statewide median income for Massachusetts AND the percentage of people 25 years and over with a bachelors degree or higher does not exceed 25 percent*; *or*
* 25 percent or more of the residents are lacking English language proficiency *AND the median annual household income does not exceed 125 percent of the statewide median income for Massachusetts AND the percentage of people 25 years and over with a bachelors degree or higher does not exceed 25 percent*.

Compared to the benchmark 2010 environmental justice community criteria, the income- and education-modified criteria of this alternative policy reduces the number of block groups qualifying as an environmental justice community by `r {sum(blkgrp_shp_WGS84$LOSEJ_inc125pctColl)}`, from `r {sum(blkgrp_shp_WGS84$CRITCNT2010 > 0)}` to `r {sum(blkgrp_shp_WGS84$CRITCNT_inc125pctColl > 0)}` in total. Under this Environmental Justice Alternative Policy D, `r {round(sum(blkgrp_shp_WGS84$CRITCNT_inc125pctColl > 0) / nrow(blkgrp_shp_WGS84) * 100,1)}`% of Massachusetts block groups are classified as environmental justice communities. This classification can be visualized in Figure \@ref(fig:treemapAltD) below, which shows the percentage of block groups that fall into each category or combination of environmental justice criteria.

```{r treemapAltD, echo=FALSE, message=FALSE, warning=FALSE, fig.align = "center", fig.cap="Tree map of block groups classified as environmental justice by Alternative EJ Policy D - Modify Minority and English Isolation Criteria by Income AND Education."}
# Create a tree map instead with treemapify
# library(treemapify)
# First, you need to generate a summary df with one value per observation or EJ criterion
blkgrp_shp_WGS84 %>% 
  mutate(EJCRIT_inc125pctColl = case_when(EJCRIT_inc125pctColl == "E" ~ "English Isolation",
                                   EJCRIT_inc125pctColl == "E  M" ~ "Minority and English Isolation", 
                                   EJCRIT_inc125pctColl == "E I" ~ "Income and English Isolation",
                                   EJCRIT_inc125pctColl == "E I M" ~ "Minority, Income, and English Isolation",
                                   EJCRIT_inc125pctColl == "I" ~ "Income",
                                   EJCRIT_inc125pctColl == "I M" ~ "Minority and Income",
                                   EJCRIT_inc125pctColl == "M" ~ "Minority",
                                   EJCRIT_inc125pctColl == "" ~ "Not EJ")) %>% 
  group_by(EJCRIT_inc125pctColl) %>% 
  summarize(blkgrps = n()) %>% 
  mutate(percent = blkgrps/sum(blkgrps)*100) %>% 
  ggplot(aes(area = percent, fill=EJCRIT_inc125pctColl, label = paste(EJCRIT_inc125pctColl,paste(round(percent,1),"%"),sep="\n"))) + 
  scale_fill_manual(name = "Environmental Justice Criteria", 
                    values = c("lightskyblue","lawngreen","cyan","gold","deeppink","darkgreen","navyblue","gray88")) +
  labs(title = "Massachusetts Census Blockgroups Meeting Alternative EJ Policy D Criteria",
       caption = "Based on Alternative EJ Policy D and ACS 5yr 2017 Blockgroup Estimates") +
  geom_treemap() + 
  geom_treemap_text(color = "white", place = "centre", grow = TRUE)
```

These classifications can be visualized and explored geographically in the interactive map below. Click on the search tool (white box in upper left of map) and type in a town or city name to zoom to that municipality. Click on individual block groups to see more detailed demographics. Click on the layers button (white box in the upper right of the map) and activate the Lost EJ Policy D layer in the map to see the block groups that are delisted (i.e. "Lost") as environmental justice communities under Poliicy D, shown as light red or salmon colored polygons. 

```{r mapAltD, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Map of block groups classified as environmental justice by Alternative Environmental Justice Policy D."}
# How many lost EJ blockgroups occur within blue communities?
lost125_50 <- blkgrp_shp_WGS84 %>% 
  as.data.frame() %>% 
  group_by(TOWN) %>% 
  summarize(lost = sum(LOSEJ_inc125pctColl)) %>% 
  left_join(macosub_shp_WGS84,.,by=c("TOWN" = "TOWN"))

# Create map of Alternative EJ Policy D polygons across state
# create df without "" in EJCRIT2010 for setting up color palette
blkgrp_shp_WGS84_naEJCRIT_inc125pctColl <- blkgrp_shp_WGS84 %>%
  select(NAME,TOWN,EJCRIT_inc125pctColl,pctminority,medhhincE,pctlimitenghh,pctCollege) %>% 
  filter(EJCRIT_inc125pctColl !="")

# create a polygon layer of blockgroups that are dropped out of EJ status by criteria change
policyDpolygons <- blkgrp_shp_WGS84 %>% 
  filter(LOSEJ_inc125pctColl == TRUE) %>% 
  select(TOWN)
  

# Create popup. NOTE THAT POPUP AND PALETTE MUST BE CREATED FROM EXACT SAME DF AS USED IN leaflet() MAP. OTHERWISE, POPUPS WILL SHOW UP IN THE WRONG LOCATION!
EJDPopUp <- paste0(blkgrp_shp_WGS84_naEJCRIT_inc125pctColl$NAME, "<br/>",
                   "<b>Town:</b> ", blkgrp_shp_WGS84_naEJCRIT_inc125pctColl$TOWN, "<br/>",
                   "<b>Policy C EJ Criteria Met:</b> ", blkgrp_shp_WGS84_naEJCRIT_inc125pctColl$EJCRIT_inc125pctColl, "<br/>",
                   "<b>ACS 5yr 2017</b> ", "<br/>",
                   "<b>% Minority: </b>", round(blkgrp_shp_WGS84_naEJCRIT_inc125pctColl$pctminority,1),"<br/>",
                   "<b>Median Household Income: </b>", paste0("$",round(blkgrp_shp_WGS84_naEJCRIT_inc125pctColl$medhhincE)),"<br/>",
                   "<b>% Limited English Households: </b>", round(blkgrp_shp_WGS84_naEJCRIT_inc125pctColl$pctlimitenghh,1),"<br/>",
                   "<b>% College Degree: </b>", round(blkgrp_shp_WGS84_naEJCRIT_inc125pctColl$pctCollege,1))

# draw the map of EJ Polygons according to EJ Policy D. Faded red/salmong colored polygons represent blockgroups that lose EJ status when the income criteria modifies minority and English isolation. 
leaflet() %>% 
  addProviderTiles(providers$Stamen.TonerLite) %>% 
  addPolygons(data = town_shp_WGS84,
              weight = 0.7,
              opacity = 1,
              color = "gray",
              fillOpacity = 0,
              label=~TOWN, popup=~TOWN, group='muni') %>% 
  addPolygons(data = macosub_shp_WGS84,
              weight = 2,
              opacity = 1,
              color = "blue",
              fillOpacity = 0.1,
              fillColor = "blue",
              label = ~TOWN,popup = TownMedIncPopUp,
              group = "Top 20% Median Household Income") %>% 
  addPolygons(data = policyDpolygons,
              fillColor = "red",
              weight = 1,
              opacity = 1,
              color = "yellow",
              dashArray = 3,
              fillOpacity = 0.3,
              label=~TOWN,
              popup = "Lost EJ Under Policy D",
              group = "Lost EJ Policy D") %>% 
  addPolygons(data = blkgrp_shp_WGS84_naEJCRIT_inc125pctColl,
              fillColor = ~EJ2010pal(EJCRIT_inc125pctColl), 
              weight = 0.5,
              opacity = 0.7,
              color = "white",
              dashArray = 3,
              fillOpacity = 0.7,
              label=~TOWN, 
              popup = EJDPopUp,
              highlight = highlightOptions(
                weight = 5,
                color = "#666",
                dashArray = "",
                fillOpacity = 0.7,
                bringToFront = TRUE)) %>% 
  addLegend(colors = EJcolsHex, 
            labels = c("Minority","Income","English Isolation","Minority and Income","Minority and English Isolation","Income and English Isolation","Minority, Income, and English Isolation"), title = "Environmental Justice Populations", position = "bottomleft") %>% 
  setView(lng = -72, 42.1, zoom = 8) %>% 
  # addMiniMap() %>% 
  addScaleBar(position = "bottomright") %>% 
  addSearchFeatures(targetGroups  = 'muni', 
                    options = searchFeaturesOptions(zoom=14, openPopup=TRUE, hideMarkerOnCollapse=T)) %>% 
  addLayersControl(
    overlayGroups = c("Lost EJ Policy D","Top 20% Median Household Income"),
    options = layersControlOptions(collapsed = TRUE)
  ) %>% 
  hideGroup(c("Top 20% Median Household Income", "Lost EJ Policy D"))
```

Under this proposal, block groups within I-495, especially the inner communities around Boston see the largest loss of EJ status. Another way to evaluate the impact of Policy D is to look at how many "lost" environmental justice classifications occur within wealthier municipalities. Under Policy D, `r {sum(lost125_50$lost)}` (`r {round(sum(lost125_50$lost) / sum(blkgrp_shp_WGS84$LOSEJ_inc125pctColl) * 100,1)}`%) of the `r {sum(blkgrp_shp_WGS84$LOSEJ_inc125pctColl)}` lost environmental justice classifications occur in these highest income municipalities. The latter addresses the issue of environmental justice designations that occur in communities not typically associated with environmental injustices. Activate the Top 20% Median Household Income layer in the map to see the municipalities in the top fifth of median household incomes (i.e. median household incomes ranging from \$105,169 to \$204,018), which are shaded in blue. 
