# Addendum 5 - Minority40 or Minority25 and Income150 {-}

This Addendum 5 option uses a threshold of 40% or higher for the minority criterion and separately adds an income-modified minority criterion. For the latter, block groups qualify for environmental justice status if they are 25% or greater minority and the median income of the block group does not exceed 150% of the statewide median (i.e. \$111,250.50).

Under this Addendum 5 policy alternative, environmental justice communities are defined as neighborhoods (U.S. Census Bureau census block groups) that meet *one or more* of the following criteria: 

* The median annual household income is at or below 65 percent of the statewide median income for Massachusetts; *or*
* *40 percent or more of the residents are minority*; *or*
* *25 percent or more of the residents are minority AND the median annual household income does not exceed 150 percent of the statewide median income for Massachusetts*; *or*
* 25 percent or more of the residents are lacking English language proficiency

Compared to the benchmark 2010 environmental justice community criteria, the minority- and income-modified criteria of this alternative policy reduces the number of block groups qualifying as an environmental justice community by `r {sum(blkgrp_shp_WGS84$LOSEJ_m40_m25inc150)}`, from `r {sum(blkgrp_shp_WGS84$CRITCNT2010 > 0)}` to `r {sum(blkgrp_shp_WGS84$CRITCNT_m40_m25inc150 > 0)}` in total. Under this Environmental Justice Addendum 5 alternative policy, `r {round(sum(blkgrp_shp_WGS84$CRITCNT_m40_m25inc150 > 0) / nrow(blkgrp_shp_WGS84) * 100,1)}`% of Massachusetts block groups are classified as environmental justice communities. This classification can be visualized in Figure \@ref(fig:treemapAd5) below, which shows the percentage of block groups that fall into each category or combination of environmental justice criteria.

```{r treemapAd5, echo=FALSE, message=FALSE, warning=FALSE, fig.align = "center", fig.cap="Tree map of block groups classified as environmental justice by Addendum 5 Policy Alternative - Minority40 or Minority25 and Income150."}
# Create a tree map instead with treemapify
# library(treemapify)
# First, you need to generate a summary df with one value per observation or EJ criterion
blkgrp_shp_WGS84 %>% 
  mutate(EJCRIT_m40_m25inc150 = case_when(EJCRIT_m40_m25inc150 == "E" ~ "English Isolation",
                                  EJCRIT_m40_m25inc150 == "E  M" ~ "Minority and English Isolation", 
                                  EJCRIT_m40_m25inc150 == "E I" ~ "Income and English Isolation",
                                  EJCRIT_m40_m25inc150 == "E I M" ~ "Minority, Income, and English Isolation",
                                  EJCRIT_m40_m25inc150 == "I" ~ "Income",
                                  EJCRIT_m40_m25inc150 == "I M" ~ "Minority and Income",
                                  EJCRIT_m40_m25inc150 == "M" ~ "Minority",
                                  EJCRIT_m40_m25inc150 == "" ~ "Not EJ")) %>% 
  group_by(EJCRIT_m40_m25inc150) %>% 
  summarize(blkgrps = n()) %>% 
  mutate(percent = blkgrps/sum(blkgrps)*100) %>% 
  ggplot(aes(area = percent, fill=EJCRIT_m40_m25inc150, label = paste(EJCRIT_m40_m25inc150,paste(round(percent,1),"%"),sep="\n"))) + 
  scale_fill_manual(name = "Environmental Justice Criteria", 
                    values = c("lightskyblue","lawngreen","cyan","gold","deeppink","darkgreen","navyblue","gray88")) +
  labs(title = "Massachusetts Census Blockgroups Meeting Addendum 5 Alternative EJ Policy Criteria",
       caption = "Based on ACS 5yr 2017 Blockgroup Estimates") +
  geom_treemap() + 
  geom_treemap_text(color = "white", place = "centre", grow = TRUE)
```

These classifications can be visualized and explored geographically in the interactive map below. Click on the search tool (white box in upper left of map) and type in a town or city name to zoom to that municipality. Click on individual block groups to see more detailed demographics. Click on the layer button (white box in the upper right of the map) and activate the Lost EJ Ad5 Policy layer in the map to see the block groups that are delisted (i.e. "Lost") as environmental justice communities under Poliicy Addendum 5, shown as light red or salmon colored polygons.

```{r mapAd5, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Map of block groups classified as environmental justice by Addendum 5 Alternative Environmental Justice Policy."}
# How many lost EJ blockgroups occur within blue communities?
lostm40_m25inc150 <- blkgrp_shp_WGS84 %>% 
  as.data.frame() %>% 
  group_by(TOWN) %>% 
  summarize(lost = sum(LOSEJ_m40_m25inc150)) %>% 
  left_join(macosub_shp_WGS84,.,by=c("TOWN" = "TOWN"))

# Create map of Addendum 5 Alternative EJ Policy polygons across state
# create df without "" in EJCRIT2010 for setting up color palette
blkgrp_shp_WGS84_naEJCRIT_m40_m25inc150 <- blkgrp_shp_WGS84 %>%
  select(NAME,TOWN,EJCRIT_m40_m25inc150,pctminority,medhhincE,pctlimitenghh) %>% 
  filter(EJCRIT_m40_m25inc150 !="")

# create a polygon layer of blockgroups that are dropped out of EJ status by criteria change
policyAd5polygons <- blkgrp_shp_WGS84 %>% 
  filter(LOSEJ_m40_m25inc150 == TRUE) %>% 
  select(TOWN)
  

# Create popup. NOTE THAT POPUP AND PALETTE MUST BE CREATED FROM EXACT SAME DF AS USED IN leaflet() MAP. OTHERWISE, POPUPS WILL SHOW UP IN THE WRONG LOCATION!
EJAd5PopUp <- paste0(blkgrp_shp_WGS84_naEJCRIT_m40_m25inc150$NAME, "<br/>",
                   "<b>Town:</b> ", blkgrp_shp_WGS84_naEJCRIT_m40_m25inc150$TOWN, "<br/>",
                   "<b>Ad5 Policy EJ Criteria Met:</b> ", blkgrp_shp_WGS84_naEJCRIT_m40_m25inc150$EJCRIT_m30inc150, "<br/>",
                   "<b>ACS 5yr 2017</b> ", "<br/>",
                   "<b>% Minority: </b>", round(blkgrp_shp_WGS84_naEJCRIT_m40_m25inc150$pctminority,1),"<br/>",
                   "<b>Median Household Income: </b>", paste0("$",round(blkgrp_shp_WGS84_naEJCRIT_m40_m25inc150$medhhincE)),"<br/>",
                   "<b>% Limited English Households: </b>", round(blkgrp_shp_WGS84_naEJCRIT_m40_m25inc150$pctlimitenghh,1))

# draw the map of EJ Polygons according to Addendum 5 EJ Policy. Faded red/salmong colored polygons represent blockgroups that lose EJ status when the income criteria modifies minority and English isolation. 
leaflet() %>% 
  addProviderTiles(providers$Stamen.TonerLite) %>% 
  addPolygons(data = town_shp_WGS84,
              weight = 0.7,
              opacity = 1,
              color = "gray",
              fillOpacity = 0,
              label=~TOWN, popup=~TOWN, group='muni') %>% 
  addPolygons(data = macosub_shp_WGS84,
              weight = 2,
              opacity = 1,
              color = "blue",
              fillOpacity = 0.1,
              fillColor = "blue",
              label = ~TOWN,popup = TownMedIncPopUp,
              group = "Top 20% Median Household Income") %>% 
  addPolygons(data = policyAd5polygons,
              fillColor = "red",
              weight = 1,
              opacity = 1,
              color = "yellow",
              dashArray = 3,
              fillOpacity = 0.3,
              label=~TOWN,
              popup = "Lost EJ Under Addendum 5 Policy",
              group = "Lost EJ Policy Ad5") %>% 
  addPolygons(data = blkgrp_shp_WGS84_naEJCRIT_m40_m25inc150,
              fillColor = ~EJ2010pal(EJCRIT_m40_m25inc150), 
              weight = 0.5,
              opacity = 0.7,
              color = "white",
              dashArray = 3,
              fillOpacity = 0.7,
              label=~TOWN, 
              popup = EJAd5PopUp,
              highlight = highlightOptions(
                weight = 5,
                color = "#666",
                dashArray = "",
                fillOpacity = 0.7,
                bringToFront = TRUE)) %>% 
  addLegend(colors = EJcolsHex, 
            labels = c("Minority","Income","English Isolation","Minority and Income","Minority and English Isolation","Income and English Isolation","Minority, Income, and English Isolation"), title = "Environmental Justice Populations", position = "bottomleft") %>% 
  setView(lng = -72, 42.1, zoom = 8) %>% 
  # addMiniMap() %>% 
  addScaleBar(position = "bottomright") %>% 
  addSearchFeatures(targetGroups  = 'muni', 
                    options = searchFeaturesOptions(zoom=14, openPopup=TRUE, hideMarkerOnCollapse=T)) %>% 
  addLayersControl(
    overlayGroups = c("Lost EJ Policy Ad5","Top 20% Median Household Income"),
    options = layersControlOptions(collapsed = TRUE)
  ) %>% 
  hideGroup(c("Top 20% Median Household Income", "Lost EJ Policy Ad5"))
```

Another way to evaluate the impact of Addendum 5 Policy is to look at how many "lost" environmental justice classifications occur within wealthier municipalities. Under Addendum 5 Policy, `r {sum(lostm40_m25inc150$lost)}` (`r {round(sum(lostm40_m25inc150$lost) / sum(blkgrp_shp_WGS84$LOSEJ_m40_m25inc150) * 100,1)}`%) of the `r {sum(blkgrp_shp_WGS84$LOSEJ_m40_m25inc150)}` lost environmental justice classifications occur in these highest income municipalities. The latter addresses the issue of environmental justice designations that occur in communities not typically associated with environmental injustices. Activate the Top 20% Median Household Income layer in the map to see the municipalities in the top fifth of median household incomes (i.e. median household incomes ranging from \$105,169 to \$204,018), which are shaded in blue. 