# Environmental Justice Alternative Policy J - Minority40 or Minority25 and MuniIncome150

This Environmental Justice Alternative Policy J option uses a threshold of 40% or higher for the minority criterion and separately adds an income-modified minority criterion. For the latter, a block group qualifies for environmental justice status if it has a population that is 25% or greater minority and the block group is *within a city or town* whose median household income does not exceed 150% of the statewide median (i.e. \$111,250.50).

Under this Environmental Justice Alternative Policy J policy alternative, environmental justice communities are defined as neighborhoods (U.S. Census Bureau census block groups) that meet *one or more* of the following criteria: 

* The median annual household income is at or below 65 percent of the statewide median income for Massachusetts; *or*
* *40 percent or more of the residents are minority*; *or*
* *25 percent or more of the residents are minority AND the block group is within a city or town whose median annual household income does not exceed 150 percent of the statewide median income for Massachusetts*; *or*
* 25 percent or more of the residents are lacking English language proficiency

Compared to the benchmark 2010 environmental justice community criteria, the minority- and income-modified criteria of this alternative policy reduces the number of block groups qualifying as an environmental justice community by `r {sum(blkgrp_shp_WGS84$LOSEJ_m40_m25Muninc150)}`, from `r {sum(blkgrp_shp_WGS84$CRITCNT2010 > 0)}` to `r {sum(blkgrp_shp_WGS84$CRITCNT_m40_m25Muninc150 > 0)}` in total. Under this Environmental Justice Environmental Justice Alternative Policy J, `r {round(sum(blkgrp_shp_WGS84$CRITCNT_m40_m25Muninc150 > 0) / nrow(blkgrp_shp_WGS84) * 100,1)}`% of Massachusetts block groups are classified as environmental justice communities. This classification can be visualized in Figure \@ref(fig:treemapAltJ) below, which shows the percentage of block groups that fall into each category or combination of environmental justice criteria.

```{r treemapAltJ, echo=FALSE, message=FALSE, warning=FALSE, fig.align = "center", fig.cap="Tree map of block groups classified as environmental justice by Environmental Justice Alternative Policy J - Minority40 or Minority25 and MuniIncome150."}
# Create a tree map instead with treemapify
# library(treemapify)
# First, you need to generate a summary df with one value per observation or EJ criterion
blkgrp_shp_WGS84 %>% 
  mutate(EJCRIT_m40_m25Muninc150 = case_when(EJCRIT_m40_m25Muninc150 == "E" ~ "English Isolation",
                                  EJCRIT_m40_m25Muninc150 == "E  M" ~ "Minority and English Isolation", 
                                  EJCRIT_m40_m25Muninc150 == "E I" ~ "Income and English Isolation",
                                  EJCRIT_m40_m25Muninc150 == "E I M" ~ "Minority, Income, and English Isolation",
                                  EJCRIT_m40_m25Muninc150 == "I" ~ "Income",
                                  EJCRIT_m40_m25Muninc150 == "I M" ~ "Minority and Income",
                                  EJCRIT_m40_m25Muninc150 == "M" ~ "Minority",
                                  EJCRIT_m40_m25Muninc150 == "" ~ "Not EJ")) %>% 
  group_by(EJCRIT_m40_m25Muninc150) %>% 
  summarize(blkgrps = n()) %>% 
  mutate(percent = blkgrps/sum(blkgrps)*100) %>% 
  ggplot(aes(area = percent, fill=EJCRIT_m40_m25Muninc150, label = paste(EJCRIT_m40_m25Muninc150,paste(round(percent,1),"%"),sep="\n"))) + 
  scale_fill_manual(name = "Environmental Justice Criteria", 
                    values = c("lightskyblue","lawngreen","cyan","gold","deeppink","darkgreen","navyblue","gray88")) +
  labs(title = "Massachusetts Census Blockgroups Meeting Environmental Justice Alternative Policy J Criteria",
       caption = "Based on ACS 5yr 2017 Blockgroup Estimates") +
  geom_treemap() + 
  geom_treemap_text(color = "white", place = "centre", grow = TRUE)
```

These classifications can be visualized and explored geographically in the interactive map below. Click on the search tool (white box in upper left of map) and type in a town or city name to zoom to that municipality. Click on individual block groups to see more detailed demographics. Click on the layer button (white box in the upper right of the map) and activate the Lost EJ Policy J layer in the map to see the block groups that are delisted (i.e. "Lost") as environmental justice communities under Poliicy Environmental Justice Alternative Policy J, shown as light red or salmon colored polygons.

```{r mapAltJ, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Map of block groups classified as environmental justice by Environmental Justice Alternative Policy J."}
# How many lost EJ blockgroups occur within blue communities?
lostm40_m25Muninc150 <- blkgrp_shp_WGS84 %>% 
  as.data.frame() %>% 
  group_by(TOWN) %>% 
  summarize(lost = sum(LOSEJ_m40_m25Muninc150)) %>% 
  left_join(macosub_shp_WGS84,.,by=c("TOWN" = "TOWN"))

# Create map of Addendum 5 Alternative EJ Policy polygons across state
# create df without "" in EJCRIT2010 for setting up color palette
blkgrp_shp_WGS84_naEJCRIT_m40_m25Muninc150 <- blkgrp_shp_WGS84 %>%
  select(NAME,TOWN,EJCRIT_m40_m25Muninc150,pctminority,medhhincE,pctlimitenghh) %>% 
  filter(EJCRIT_m40_m25Muninc150 !="")

# create a polygon layer of blockgroups that are dropped out of EJ status by criteria change
policyAltJpolygons <- blkgrp_shp_WGS84 %>% 
  filter(LOSEJ_m40_m25Muninc150 == TRUE) %>% 
  select(TOWN)
  

# Create popup. NOTE THAT POPUP AND PALETTE MUST BE CREATED FROM EXACT SAME DF AS USED IN leaflet() MAP. OTHERWISE, POPUPS WILL SHOW UP IN THE WRONG LOCATION!
EJAltJPopUp <- paste0(blkgrp_shp_WGS84_naEJCRIT_m40_m25Muninc150$NAME, "<br/>",
                   "<b>Town:</b> ", blkgrp_shp_WGS84_naEJCRIT_m40_m25Muninc150$TOWN, "<br/>",
                   "<b>Policy J EJ Criteria Met:</b> ", blkgrp_shp_WGS84_naEJCRIT_m40_m25Muninc150$EJCRIT_m40_m25Muninc150, "<br/>",
                   "<b>ACS 5yr 2017</b> ", "<br/>",
                   "<b>% Minority: </b>", round(blkgrp_shp_WGS84_naEJCRIT_m40_m25Muninc150$pctminority,1),"<br/>",
                   "<b>Median Household Income: </b>", paste0("$",round(blkgrp_shp_WGS84_naEJCRIT_m40_m25Muninc150$medhhincE)),"<br/>",
                   "<b>% Limited English Households: </b>", round(blkgrp_shp_WGS84_naEJCRIT_m40_m25Muninc150$pctlimitenghh,1))

# draw the map of EJ Polygons according to Environmental Justice Alternative Policy J EJ Policy. Faded red/salmong colored polygons represent blockgroups that lose EJ status when the income criteria modifies minority and English isolation. 
leaflet() %>% 
  addProviderTiles(providers$Stamen.TonerLite) %>% 
  addPolygons(data = town_shp_WGS84,
              weight = 0.7,
              opacity = 1,
              color = "gray",
              fillOpacity = 0,
              label=~TOWN, popup=~TOWN, group='muni') %>% 
  addPolygons(data = macosub_shp_WGS84,
              weight = 2,
              opacity = 1,
              color = "blue",
              fillOpacity = 0.1,
              fillColor = "blue",
              label = ~TOWN,popup = TownMedIncPopUp,
              group = "Top 20% Median Household Income") %>% 
  addPolygons(data = policyAltJpolygons,
              fillColor = "red",
              weight = 1,
              opacity = 1,
              color = "yellow",
              dashArray = 3,
              fillOpacity = 0.3,
              label=~TOWN,
              popup = "Lost EJ Under Environmental Justice Alternative Policy J",
              group = "Lost EJ Policy J") %>% 
  addPolygons(data = blkgrp_shp_WGS84_naEJCRIT_m40_m25Muninc150,
              fillColor = ~EJ2010pal(EJCRIT_m40_m25Muninc150), 
              weight = 0.5,
              opacity = 0.7,
              color = "white",
              dashArray = 3,
              fillOpacity = 0.7,
              label=~TOWN, 
              popup = EJAltJPopUp,
              highlight = highlightOptions(
                weight = 5,
                color = "#666",
                dashArray = "",
                fillOpacity = 0.7,
                bringToFront = TRUE)) %>% 
  addLegend(colors = EJcolsHex, 
            labels = c("Minority","Income","English Isolation","Minority and Income","Minority and English Isolation","Income and English Isolation","Minority, Income, and English Isolation"), title = "Environmental Justice Populations", position = "bottomleft") %>% 
  setView(lng = -72, 42.1, zoom = 8) %>% 
  # addMiniMap() %>% 
  addScaleBar(position = "bottomright") %>% 
  addSearchFeatures(targetGroups  = 'muni', 
                    options = searchFeaturesOptions(zoom=14, openPopup=TRUE, hideMarkerOnCollapse=T)) %>% 
  addLayersControl(
    overlayGroups = c("Lost EJ Policy J","Top 20% Median Household Income"),
    options = layersControlOptions(collapsed = TRUE)
  ) %>% 
  hideGroup(c("Top 20% Median Household Income", "Lost EJ Policy J"))
```

Another way to evaluate the impact of Environmental Justice Alternative Policy J is to look at how many "lost" environmental justice classifications occur within wealthier municipalities. Under Environmental Justice Alternative Policy J, `r {sum(lostm40_m25Muninc150$lost)}` (`r {round(sum(lostm40_m25Muninc150$lost) / sum(blkgrp_shp_WGS84$LOSEJ_m40_m25Muninc150) * 100,1)}`%) of the `r {sum(blkgrp_shp_WGS84$LOSEJ_m40_m25Muninc150)}` lost environmental justice classifications occur in these highest income municipalities. The latter addresses the issue of environmental justice designations that occur in communities not typically associated with environmental injustices. Activate the Top 20% Median Household Income layer in the map to see the municipalities in the top fifth of median household incomes (i.e. median household incomes ranging from \$105,169 to \$204,018), which are shaded in blue. 